A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2023-01-06 paper/paper.tex
paper: minor polish

\title{A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{IRIF/Inria Paris/Université Paris-Cité}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université Paris-Cité}
}
\date{}

@@

We exploit the connection between iterated unary parametricity and augmented simplicial sets, and between iterated binary parametricity and cubical sets to present a uniform construction of both semi-simplicial semi-cubical sets obtained by applying simple rules iterating parametricity. In contrast to the presheaf presentation of semi-simplicial and semi-cubical sets, our uniform construction, which we call \emph{$\nu$-sets}, is \emph{indexed}. That is, instead of having the set of semi-simplices or semi-cubes in dimension $n+1$
\emph{fibered} over the set of semi-simplices or semi-cubes in dimension $n$, we consider families of semi-simplices and semi-cubes indexed over their faces.

@@

Since the construction shown in the previous sections are by induction on $n$, and dependencies are on lower $n$ and $p < n$, one would imagine formalizing this using well-founded induction in the calculus of inductive constructions. We initially tried this approach: we had terms dependent on the proofs of the case distinction of $n' \leq n$ implies $n' < n$ or $n' = n$, and these proofs were didn't have definitional computational rule; it was necessary to reason propositionally on the computational property of the case distinction, and it even turned out to be unmanageable. Hence, we chose a different route: in practice, since $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$, we only need to keep track of three ``levels'', and we built separate data structures for the levels, with dependencies. More concretely, we build the ten definitions shown in the tables by induction, and this is part of the definition of a larger record. Other fields in the record are necessary because we need copies of $\framep$, $\layer$, $\filler$ at levels $n - 1$ and $n - 2$, $\restrf$, $\restrl$, and $\restrc$ at level $n - 1$, and equations to recall the definitions of these objects at lower levels.

The entire construction relies on inequalities over natural numbers, and we have three different definitions of $\leq$ addressing different concerns, and conversions between them. First, we have what we call the ``recursive variant'', phrased as:

>>>>

\title{A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
}
\date{}

@@

We exploit the connection between iterated unary parametricity and augmented simplicial sets, and between iterated binary parametricity and cubical sets to present a uniform construction of both semi-simplicial and semi-cubical sets obtained by applying simple rules iterating parametricity. In contrast to the presheaf presentation of semi-simplicial and semi-cubical sets, our uniform construction, which we call \emph{$\nu$-sets}, is \emph{indexed}. That is, instead of having the set of semi-simplices or semi-cubes in dimension $n+1$
\emph{fibered} over the set of semi-simplices or semi-cubes in dimension $n$, we consider families of semi-simplices and semi-cubes indexed over their faces.

@@

Since the construction shown in the previous sections are by induction on $n$, and dependencies are on lower $n$ and $p < n$, one would imagine formalizing this using well-founded induction in the calculus of inductive constructions. We initially tried this approach: we had terms dependent on the proofs of the case distinction of $n' \leq n$ implies $n' < n$ or $n' = n$, and these proofs didn't have definitional computational rule; it started to be necessary to reason propositionally on the computational property of the case distinction, and it eventually turned out to be unmanageable. Hence, we chose a different route: in practice, since $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$, we only need to keep track of three ``levels'', and we built separate data structures for the levels, with dependencies. More concretely, we build the ten definitions shown in the tables by induction, and this is part of the definition of a larger record. Other fields in the record are necessary because we need copies of $\framep$, $\layer$, $\filler$ at levels $n - 1$ and $n - 2$, $\restrf$, $\restrl$, and $\restrc$ at level $n - 1$, and equations to recall the definitions of these objects at lower levels.

The entire construction relies on inequalities over natural numbers, and we have three different definitions of $\leq$ addressing two different concerns, and conversions between them. First, we have what we call the ``recursive variant'', phrased as:
